---
title: "casestudy"
author: "Daniel M. Turner"
date: "December 17, 2018"
output: html_document
---

0.
```{r}
##load data
beer <- read.csv("C:\\Users\\DMT\\Documents\\masters\\semester1\\dds402\\CaseStudy_2_2_2_2_2_2\\Beers.csv", header = TRUE, sep = ",")

brewery <- read.csv("C:\\Users\\DMT\\Documents\\masters\\semester1\\dds402\\CaseStudy_2_2_2_2_2_2\\Breweries.csv", header = TRUE, sep = ",")

```


1.	How many breweries are present in each state?
```{r}
##count by state

brewsByState = summary(brewery$State)


```

2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r}
##data merge and check.clean data?

colnames(brewery)[colnames(brewery)=="Brew_ID"] <- "Brewery_id"
brewNBeer <- merge(beer, brewery, by = "Brewery_id")
head(brewNBeer)
tail(brewNBeer)


```

3.	Report the number of NA's in each column.
```{r}
##after clean, only missing bitterness? beer types?
colSums(is.na(brewNBeer))


```

4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r}
##find medians by state. Do twice? once counting NA and once not counting NA
#alcy<-tapply(brewNBeer$ABV, INDEX=brewNBeer$State, FUN=median)
#bitter<-tapply(brewNBeer$IBU, INDEX=brewNBeer$State, FUN=median)

#remove na for data
alcy<-tapply(brewNBeer$ABV[!is.na(brewNBeer$ABV)], INDEX=brewNBeer$State[!is.na(brewNBeer$ABV)], FUN=median)
bitter<-tapply(brewNBeer$IBU[!is.na(brewNBeer$IBU)], INDEX=brewNBeer$State[!is.na(brewNBeer$IBU)], FUN=median)

library(ggplot2)

df_abv <- data.frame(state=names(alcy),abv = alcy)
ggplot(data=df_abv, aes(x=state, y=abv)) +
    geom_bar(stat="identity")+ coord_flip()+
    geom_bar(stat="identity", color="blue", fill="green")+
    geom_text(aes(label=abv), hjust=-2, size=2)+
    theme_minimal()

bitter[is.na(bitter)] <- 0
df_ibu <- data.frame(state=names(bitter),ibu = bitter)
ggplot(data=df_ibu, aes(x=state, y=ibu)) +
    geom_bar(stat="identity")+ coord_flip()+
    geom_bar(stat="identity", color="blue", fill="green")+
    geom_text(aes(label=ibu), hjust=-2, size=2)+
    theme_minimal()

```

5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
##find biggest numbers. NBD
brewNBeer[which.max(brewNBeer$ABV),] #alcohol
brewNBeer[which.max(brewNBeer$IBU),] #bitterness


```

6.	Summary statistics for the ABV variable.
```{r}
##summarize numbers. box plot?
summary(brewNBeer$ABV)

```

7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r}
##plot ibu vs abv. nbd
plot(brewNBeer$ABV,brewNBeer$IBU)# ugly but a starting point


```

prep develop and present
```{r}
##1-7, describe results in common speak


```
